Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-08-25 review/response.Rmd

The tool is likely to be of interest to the R community.  Most of my comments are about the technical aspect of the content. The writing is overall good, but please see my comment under the "writing" regarding to the clarity of the text.

### Major comment

- "Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign...However, this creates difficulties for comparing the optimisers since the end bases will be symmetric to the origin. A sign flipping step is conducted..." and the "Reconciling the orientation" section
  - For 1D projections, flipping the sign is sufficient. For p-D projections of $\mathbb{R}^n$​, one may need a little more. To my knowledge most index functions are invariant to any rotations that happens within the p-subspace. This makes me wonder that in general, one may want to align two bases so that their difference does not contain any "within-subspace" components. In other words, one may want to compare representations in the Grassmannian manifold  (space of k-subspaces) instead of in the Stiefel manifold (space of orthonormal k-frames).

\textcolor{red}{Is this a comment or anything to change?}
I think the reviewer want us to comment on the sign flipping for higher order dimensions. It seems that the reviewer is suggesting that instead of comparing the bases on the space of the orthonormal k-frames it might be good to compare their representation in the Grassmannian manifold.

Grassmannian manifold: GR9k,V) is a space that parametrize all k-dimension a linear subspaces of the n-dimensional vector sace V.(Wikpedia source)

### 2021-08-25 review/response.Rmd

### Minor comment

- The software seems to be specific to projection pursuit. I wish the authors could comment more on its potential generality to other optimizations. For example, can one generalizes the tour method to visualize any parameter trace of an optimizer?

**It is feasible to generalise the tour method for diagnosing other optimisers, providing the parameter space can be simulated. In our examples, the parameter to trace (projection base ($p \times d$) ) lives in an p-sphere, which can be simulated by the `geozoo` package. **

Probably here the reviewer would like to see a kind of "recipie" about how to use the framework of the paper for other optimisers?

@@

**Changed to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis. The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."**

- "Figure 6 shows some frames from the tour plot of the same two optimisations in its original 5D space."
  - I think this is the first mention of "5D space"/5D data in the main text (apart from the caption in Figure 4), so it may be worth clarifying what the original 5D space is.

\textcolor{red}{The description of data is in the subsection simulation setup in the next section.}

It might be worth to mentioned about the 5D in the simulation sep up subsection so that it is introduced explicitly for the readers.

@@

# Reviewer 2

These experiments are interesting and so is the methodology.
The conclusions could do with beefing up
What can we learn about how to optimise the holes index for 1 and 2d?
Does this apply to other smooth indices? What about non-smooth indices?
What about indices in other dimensions?

\textcolor{red}{Is the reviewer suggesting adding more details in the first paragraph of the conclusion?}

I think the reviewer here wants to see a bit more detail on the conclusions and have provided a number of questions that could be answered in order to add some interesting points on that section.

@@

**added**

9. For the Kolmogorov index, why are you using randomly generated data instead of normal quantiles?

\textcolor{red}{TODO}

The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data. I am not sure what would be the reasoning for using normal quantiles in this index. That would imply you are assuming normality on the theoretial pdf.

### 2021-09-02 review/response.Rmd

- "Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign...However, this creates difficulties for comparing the optimisers since the end bases will be symmetric to the origin. A sign flipping step is conducted..." and the "Reconciling the orientation" section
  - For 1D projections, flipping the sign is sufficient. For p-D projections of $\mathbb{R}^n$​, one may need a little more. To my knowledge most index functions are invariant to any rotations that happens within the p-subspace. This makes me wonder that in general, one may want to align two bases so that their difference does not contain any "within-subspace" components. In other words, one may want to compare representations in the Grassmannian manifold  (space of k-subspaces) instead of in the Stiefel manifold (space of orthonormal k-frames).

\textcolor{red}{Is this a comment or anything to change?}
I think the reviewer want us to comment on the sign flipping for higher order dimensions. It seems that the reviewer is suggesting that instead of comparing the bases on the space of the orthonormal k-frames it might be good to compare their representation in the Grassmannian manifold.

Grassmannian manifold: GR9k,V) is a space that parametrize all k-dimension a linear subspaces of the n-dimensional vector sace V.(Wikpedia source)

**ANS: With p-D bases, the correction of orientation is a bit more complicated. Take 2D as an example, two bases can be aligned in the first axis but not in the second. This would means there are $2^p$ possibility to check. **

@@

**It is feasible to generalise the tour method for diagnosing other optimisers, providing the parameter space can be simulated. In our examples, the parameter to trace (projection base ($p \times d$) ) lives in an p-sphere, which can be simulated by the `geozoo` package. **

Probably here the reviewer would like to see a kind of "recipie" about how to use the framework of the paper for other optimisers?

- Figure 4 "...All the bases in PD have been flipped for easier comparison of the final bases..."
  - This is related to the major comment. In the figure, PD flipped in sign but stopped at a basis representation similar to the destination of the CRS. Flipping the sign back, does it mean that PD and CRS diverged but stopped at the similar subspace with opposite basis representations? Also, given that PD and CRS have the same starting bases and one is flipped, why wouldn't the dashed line pass the origin (gray dot) in the figure?

**Yes the reviewer is giving a decent explanation in describing the effect of flipping. The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code. This has now been corrected. **

- "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus".

  - It is not clear to me why "so". For p=2, the space of 2x2 bases is a pair of 1-D tori instead of just one (because you have two orientations for a pair of orthogonal vectors), right? For p=3, the space of 3x2 bases is SO(3), which is not a 3-D torus? Ref: https://en.wikipedia.org/wiki/Stiefel_manifold

- We change the wording of the sentence to "**a torus in the $p$-D space**"

@@

- "Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated by PCA. The centre of the circle is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."
  - It took me multiple passes to get what this means. Perhaps instead starting with something like: "We use PCA to project and visualize the parameters/bases in 2D. The centre of the 2D view is..."?

**Changed to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis. The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."**

- "Figure 6 shows some frames from the tour plot of the same two optimisations in its original 5D space."
  - I think this is the first mention of "5D space"/5D data in the main text (apart from the caption in Figure 4), so it may be worth clarifying what the original 5D space is.

**The explanation of data set is introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*. To avoid ambiguity, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...". **

@@

8. "not the case for all the indexes" missing a "." page 11

**added**

9. For the Kolmogorov index, why are you using randomly generated data instead of normal quantiles?

**The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data. I am not sure what would be the reasoning for using normal quantiles in this index. That would imply you are assuming normality on the theoretial pdf.**

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.Rmd
more changes as per the reviewers

```{r external, include = FALSE, cache = FALSE}
knitr::read_chunk(
  here::here("paper", "zhang-cook-laa-langrene-menendez.R")
)
```

@@

As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in the \pkg{geozoo} [@geozoo] package. While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation. Figure \ref{fig:toy-tour} shows some frames from the tour plot of the same two optimisations in its original 5D space.

@@

While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames. Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a $p$-D torus. For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}. The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality condition. Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous plots. The final basis found by PD and CRS are shown in a larger shape  and printed below respectively:

>>>>

```{r external, include = FALSE, cache = FALSE}
knitr::read_chunk(
  here::here("paper", "script.R")
)
```

@@

As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in the \pkg{geozoo} [@geozoo] package. While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation. Figure \ref{fig:toy-tour} shows some frames from the tour plot of the same two optimisations in its original space.

@@

While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames. Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space. For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}. The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality condition. Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous plots. The final basis found by PD and CRS are shown in a larger shape  and printed below respectively:

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.tex
more changes as per the reviewers

While the previous few examples have looked at the space of 1-frames in
a unit sphere, this section visualises the space of 2-frames. Recall
that the columns in a 2D basis are orthogonal to each other, so the
space of \(p \times 2\) bases is a \(p\)-D torus. For \(p = 3\) one
would see a classical 3D torus shape as shown by the grey points in
Figure \ref{fig:toy-torus}. The two circles of the torus can be observed
to be perpendicular to each other and this can be linked back to the
orthogonality condition. Two paths from CRS and PD are plotted on top of
the torus and coloured in green and brown respectively to match the
previous plots. The final basis found by PD and CRS are shown in a
larger shape and printed below respectively:

>>>>

While the previous few examples have looked at the space of 1-frames in
a unit sphere, this section visualises the space of 2-frames. Recall
that the columns in a 2D basis are orthogonal to each other, so the
space of \(p \times 2\) bases is a torus in the \(p\)-D space. For
\(p = 3\) one would see a classical 3D torus shape as shown by the grey
points in Figure \ref{fig:toy-torus}. The two circles of the torus can
be observed to be perpendicular to each other and this can be linked
back to the orthogonality condition. Two paths from CRS and PD are
plotted on top of the torus and coloured in green and brown respectively
to match the previous plots. The final basis found by PD and CRS are
shown in a larger shape and printed below respectively:
